---
title: "DATA607_KP_3RD_ASSGNMNT"
author: "Koohyar Pooladvand"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is a week 3 assignment, working with strings. The assignment has four different sections each will be presented below separately.

1.  Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [<https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/>], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

2.  Write code that transforms the data from an unorganized format to an organized structure extracting meaningful information

3.  Describe, in words, what these expressions will match.

4.  Construct regular expressions to match words.

## Code and assignment

In the first section of the code like, I ensure all the relevant packages are installed and libraries are loaded.

```{r, Code_initialization, echo=TRUE}
required_packages <- c("RSQLite","devtools","tidyverse","DBI","dplyr","odbc","openintro","ggplot2","psych","reshape2","knitr","markdown","shiny","R.rsp","fivethirtyeight","RCurl", "stringr","readr","glue") # Specify packages

not_installed <- required_packages[!(required_packages %in% installed.packages()[ , "Package"])]# Extract not installed packages
if(length(not_installed)==0){
  print("All required packages are installed")
} else {
  print(paste(length(not_installed), "package(s) had to be installed.")) # print the list of packages that need to be isstall
  install.packages(not_installed)
}



```

## #1 section of the 3rd week homework

The goal is to download the 173 majors and list those that contain "DATA" or "STATISTICS". Like other code, the first goal is to load the data to R, since it is URL code, we will be using the relevant line of code has been used before.

```{r Load_major_from_GitHub, echo=TRUE}
#LaodData
library('RCurl')
#for downlaoding the CSV we need to use raw code, otherwise, we will not be able to doan the file correctly
URL <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
URL_handle <- RCurl::getURL(URL)
Major_data<-data.frame(read.csv(text=URL_handle, header=TRUE,sep=","))

print("This is the size of the dataframen and let's take a look at its contents")
#dim(Major_data)
pillar::glimpse(Major_data)

```

After loading the data, we use code of strings like those that have been shared to filter for "DATA" and "STATISTICS" as well as the ones that contain both.

```{r find_string, echo=TRUE}

#using the basic function of R
#search for data
Contains_DATA_B <- grep(pattern = 'DATA', Major_data$Major, value = TRUE, ignore.case = TRUE)
#search for Science
Contains_STATISTICS_B <-grep(pattern = 'STATISTICS', Major_data$Major, value = TRUE, ignore.case = TRUE)

#search for those with either DATA or STATISTICS
print("Here is the list of the majors that have either DATA or STATISTICS")
Contains_DATA_STATISTICS_B<- grep(pattern='DATA|STATISTICS', Major_data$Major,value = TRUE, ignore.case = TRUE)
print("Result of the first way")
print(Contains_DATA_STATISTICS_B)

#In addition to grep I also will use dplyr to do the same 

#using more adacned function of dplyr
Contains_DATA<- dplyr::filter(Major_data, grepl('DATA',Major))
Contains_STATISTICS<- dplyr::filter(Major_data, grepl('STATISTICS',Major))
Contains_DATA_STATISTICS<- dplyr::filter(Major_data, grepl('STATISTICS|DATA',Major))
#Contains_DATA_STATISTICS<- dplyr::filter(Major_data, grepl('STATISTICS&DATA',Major))

print("Result of the 2nd way using dplyr")
print(Contains_DATA_STATISTICS)

```

## #2 Text readout and organization

the goal of this section is to write a code that transforms the data from an unorganized format to an organized structure extracting meaningful information. The example to be worked on it the following:

Data to be transformed:

[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

[9] "elderberry"   "lime"         "lychee"       "mulberry"    

[13] "olive"        "salal berry"

Into a format like this:

c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r read_out, echo=TRUE}

#To make this excersise realsitic I entered all those into a text file and them will
#read them into RStudio usin readLines 

con <- file("Data/raw_data.txt","r", blocking = FALSE) #Define local connection to file in Data folder

lines <- readLines(con, n= -1L, encoding="unknown", warn = TRUE, skipNul = FALSE) #read all lines 
singleString <- paste(lines, collapse=" ")

close(con) #close con
lines

#second way use readr cuntion from readr library 
library(readr)
mystring <- read_file("Data/raw_data.txt")
mystring

```

### Manipulate the uploaded text

In section of the code, I will be using regular expression to manipulate the data that are uplaoded.

#### First solution

```{r text_maniupulation_1, echo=TRUE}

# Extract words between quotations using regular expressions
words <- unlist(regmatches(lines, gregexpr('"([^"]*)"', lines)))
#print the results all contents were seperated by only words between qoutations 
print("Initial extracted words")
print(words)

#substitute the two qoutation 
#'"([^"]+)"' matches anything between double quotes and excludes the quotes themselves. The \\1 in the gsub function replaces the matched pattern with just the content within the quotes.

words <- gsub('"([^"]+)"', '\\1', words)

print("Words after spaces and extra qoutatuons are removed")
#Seperate all words 
print(words)

# Remove empty elements
words <- words[words != ""]
#Print the modified words
print("Words after all empty cells are removed")
print(words)

#Prepared the text as ask to be printed out 
#I will used glue from glue library to do the final print

print("Here is the final results")
glue::glue(
  'c("{combined}")',
  combined=glue::glue_collapse(
    words,
    sep='" ,"',
    last=""
  )
)|>print()

```

## #3 Describe, in words, what these expressions will match

-   (.)\\1\\1 :

-   "(.)(.)\\\\2\\\\1"

-   (..)\\1

-   "(.).\\\\1.\\\\1"

-   "(.)(.)(.).\*\\\\3\\\\2\\\\1"

| Expression                    | Explanation in words                                                                                                                                                                                                                        |
|--------------|----------------------------------------------------------|
| (.)\\1\\1                     | matches any sequence of three consecutive characters where the first and second characters are the same                                                                                                                                     |
| "(.)(.)\\\\2\\\\1"            | matches any sequence of four characters where the 1st and 4th are the same, and the 2nd and 3rd are the same but can be different or the same as the 1st and 4th.                                                                           |
| (..)\\1                       | matches any sequence of four characters where the first two characters are identical to the last two characters.                                                                                                                            |
| "(.).\\\\1.\\\\1"             | matches any sequence of five characters where, and the 1st and 3rd are the same, the 2nd is any character, the 5th is the same as the first.                                                                                                |
| "(.)(.)(.).\*\\\\3\\\\2\\\\1" | matches a sequence of characters that starts with any three characters, followed by any or not character, and finally ends with the same characters but reversely meaning from the third first characters toward the first first character. |

: Described the expression in words

## #4:Construct regular expressions to match words.

| Explanation                                                                               | Regular expressions   |
|------------------------------------------|------------------------------|
| Start and end with the same character.                                                    | \^(.).\*\\1\$         |
| Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)          | .\*(.{2}).\*\\1.\*    |
| Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.) | .\*(.).\*\\1.\*\\1.\* |

: Construct regular expressions
